---
title: "Linear regression - inference"
author: ""
date: ""
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      comment = "#>", highlight = TRUE,
                      fig.align = "center")
```

## Main ideas

- Reinforce linear model fundamentals

- Inference for $\beta_0$ and $\beta_1$

## Packages

```{r packages, include=FALSE}
library(tidyverse)
library(fivethirtyeight)
library(broom)
```

## Data

```{r load_data}
glimpse(candy_rankings)
```

# Notes

## Sugar model

Let's modify the sugar variable again for easier interpretation.

```{r mutate_sugar}
candy_rankings <- candy_rankings %>%
  mutate(sugarpercent100 = sugarpercent * 100)
```

What units are the explanatory and response variable in? 

```{r sugar_model}
sugar_model <- lm(winpercent ~ sugarpercent100, data = candy_rankings)
tidy(sugar_model)
```

$$\widehat{WinPercent} = 44.6 + 0.12~SugarPercentile$$

**Interpretation**:

Why does having the sugar variable on a scale from 0 to 100 lead to a better 
interpretation than if we had the scale go from 0 to 1?

Another way to interpret linear regression coefficients is by looking at 
a *one standard deviation* increase rather than a one unit increase. 

**Question:** Why might you want to look at this instead of a one unit increase?

You can find a one standard deviation increase by multiplying the coefficient 
for that term by the standard deviation.

What is the standard deviation of the sugar percentile variable?

```{r sd_sugar}

```

Now, let's multiply it by the coefficient here.

```{r sd_increase}

```


**Interpretation**:

**Question**: Would you always want to use this sort of interpretation? Does it 
make sense to do a one standard deviation increase interpretation when working 
with a dummy variable?

## Assessing the model fit with $R^2$

- One approach to quantify the quality of the fit of a linear model is $R^2$.

- It tells us what percentage of the variability in the response variable is 
  explained by the model. The remainder of the variability is unexplained.

- $R^2$ is also known as the coefficient of determination.

-**Question**: What does "explained variability in the response variable" mean?

### Calculating $R^2$

```{r r_squared}


```

Roughly 5% of the variability in the percent of time a candy bar wins can be 
explained by the sugar percentile.

Is this a high or low $R^2$ value?

**Question**: What is the correlation between the sugar percentile and 
winning percentage variables? How does this number relate to the $R^2$ value 
here?

```{r correlation}


```

### $R^2$ principles

- We can write explained variation using the following ratio of sums of squares:

$$ R^2 =  1 - \left( \frac{ SS\_{Error} }{ SS\_{Total} } \right) $$

where $SS_{Error}$ is the sum of squared residuals and $SS_{Total}$ is the total 
variance in the response variable.

Next class, when we talk about multiple regression, we will discuss another 
measure of model fit called Adjusted $R^2$ that is preferable for models with 
many explanatory variables.

## CLT-based inference for regression coefficients

- Population model:

$$ y = \beta_0 + \beta_1~x_1 + \epsilon $$

- Sample model that we use to estimate the population model:
  
$$ \hat{y} = b_0 + b_1~x_1  $$

Similar to other sample statistics (mean, proportion, etc) there is variability 
in our estimates of the slope and intercept. 

- Do we have convincing evidence that the true linear model has a non-zero 
  slope?

- What is a confidence interval for the population regression coefficient?

Let's return to our sugar content model:
```{r visualizingmodel}
ggplot(data = candy_rankings, aes(x = sugarpercent100, y = winpercent)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Do Sugary Candies Win More Often?", 
       x = "Sugar Percentile", y = "Win Percentage") + 
  theme_minimal()
```

### Tidy confidence interval

Recall our confidence interval formula:

$$\mbox{point estimate} \pm \mbox{margin of error},$$

where the margin of error is a function of the standard error and a 
multiplicative quantity derived from the point estimate's standardized
distribution.

A confidence interval for $\beta_`$ is given by

$$b_1 \pm t^*_{n-2} \times SE_{b_1}$$

Function `tidy()` can compute the interval endpoints.

```{r tidy_confint}
tidy(sugar_model, conf.int = TRUE, conf.level = 0.95)
```

A 95% confidence interval for $\beta_1$ is given by

```{r sugar_ci}


```

### Interpretation

We are 95% confident that for every additional percentile of sugar, the win 
percentage is expected to increase, on average, between 0.008 and 0.23 
percentage points.

### Hypothesis testing for $\beta_1$

Is there convincing evidence, based on our sample data, that sugar content is 
associated with winning percentage?

We can set this up as a hypothesis test.

- There is no relationship, the slope is 0.

$$H_0: \beta_1 = 0$$

- There is a relationship between sugar content and winning percentage.

$$H_A: \beta_1 \ne 0$$ 

We only reject $H_0$ in favor of $H_A$ if the data provide strong evidence
that the true slope parameter is different from zero. 

Under the null hypothesis, we can define a quantity $T$ such that

$$T = \frac{b_k - 0}{SE_{b_k}},$$

where $T$ follows a t-distribution with $n-2$ degrees of freedom. The p-values 
in the `tidy()` output represent the two-sided p-value associated with the 
observed statistic.
 
```{r slope_test}


```

Based on the result above we reject $H_0$ in favor of $H_A$. We have enough 
evidence to suggest that there is an association between sugar content and
win percentage.

## Practice

```{r brain_data}
brain <- read_table("http://users.stat.ufl.edu/~winner/data/brainhead.dat",
                    col_names = c("gender", "age_range", "head_size", 
                                  "brain_weight"))
```

A short description of the variables is given at 
http://users.stat.ufl.edu/~winner/data/brainhead.txt.

1. Create a scatterplot of `brain_weight` versus `head_size`. Comment on if
   you think a linear model is appropriate.

```{r practice_1}

```

2. Is there evidence to suggest that head size is associated with brain weight?
   Perform a statistical hypothesis test at the 0.01 significance level.
   
```{r practice_2}

```

3. Fit a linear model with brain weight and variable head size standardized.
   Interpret each of the coefficients.
   
```{r practice_3}

```
   
4. Fit a linear model with brain weight as the response and age range as the
   single predictor. Create and interpret 95% confidence interval for the
   coefficients.
   
```{r practice_4}

```

## References

1. R.J. Gladstone (1905). "A Study of the Relations of the Brain to 
   to the Size of the Head", Biometrika, Vol. 4, pp105-123